---
title: "Assignment 3"
author: "Amelia Walsh"
date: "1/27/2021"
output: html_document
---
### Project Info
Date: January 27, 2021

Link: https://github.com/15aw57/Amelia-Assignments

### Load Data
```{r}
MyData = read.csv("InputData/FallopiaData.csv")

#Check out data
str(MyData)
```

### Remove rows with ‘Total’ biomass < 60
```{r}
#Load necessary libraries
library(dplyr)
library(knitr)

#Use "filter" to remove rows from MyData
MyData = 
  MyData %>%
  filter(Total<60)
```

### Reorder the columns so that they are in the order: ‘Total’, ‘Taxon’, ‘Scenario’, ‘Nutrients’, and remove the other columns
```{r}
#Use "select" to remove unnecessary columns
MyData =
  MyData %>% 
  select(Total, Taxon, Scenario, Nutrients) %>%
  #Use arrange to reorder columns
  arrange(Total, Taxon, Scenario, Nutrients)
  
```

### Make a new column TotalG, which converts the ‘Total’ column from mg to grams AND replace Total with TotalG, and add it to the dataset.
```{r}
#Use "Mutate" to create a new column that equals 'Total'/1000
MyData =
  MyData %>%
  mutate(TotalG = Total/1000) %>%
  #Use select to remove 'Total' column
  select(-Total)
```

### Write a custom function that will take two inputs from the user: 1. a vector of data to process (e.g. column from a data.frame object) and 2. a string that defines what calculation to perform.
- if string #2 is "Average" then calculate the average value for the column named in vector #1
- if string #2 is "Sum" then calculate the sum of values for the column named in vector #1
- if string #2 is "Observations" then count the number of observed values for the column named in vector #1
- if string #2 is anything else, then output an error to the user 

```{r}
#Create a custom function with two inputs
calc  = function(vector,calctype){
  
  #If/Then function that will return the average of the vector when "Average" is input
  if(calctype == "Average"){
    return(mean(vector))
  }
  
  #Else/If function that will return the Sum of the vector when "Sum" is input
  else if(calctype == "Sum"){
    return(sum(vector))
  }
  
  #Else/If function that will return number of observations in vector when "Observations" is input
  else if(calctype == "Observations"){
    return(NROW(vector))
    
    
  }
  
  #Function that will return error message when anything other than previous commands are is input
  else
    return("Error in calctype")
    }

```

### Write some R code that uses your function to count the total number of observations in the 'Taxon' column.
```{r}
#Use custom function from prior code
calc(MyData$Taxon, "Observations")
```

### Write some R code that uses your function  to calculate the average TotalG for each of the two Nutrient concentrations.
```{r}
#Use dplyr to filter "high" nutrients from MyData
Dat2 = MyData %>%
  filter(Nutrients == "high")

#Calculate the average TotalG of high nutrients using custom function
  calc(Dat2$TotalG, "Average") 
```
```{r}
#Use dplyr to filter "low" nutrients from MyData
Dat3 = MyData %>%
  filter(Nutrients == "low")

#Calculate the average TotalG of low nutrients using custom function
  calc(Dat3$TotalG, "Average") 
```

### Write (i.e. save) the new data to a file called "WrangledData.csv" in the Output folder. 
```{r}
write.csv(MyData,"WrangledData.csv") 
```

